---
title: "P-hacking COVID-19 (c.elegans)"
author: "c.elegans"
date: "9/15/2020"
output: html_document
---

```{r global_options, include=FALSE}
# knitr::opts_chunk$set(echo=TRUE, eval=FALSE, warning=FALSE, message=FALSE)
knitr::opts_chunk$set(error=TRUE, warning=FALSE, message=FALSE)

```

```{r, message = FALSE}
library(tidyverse)
library(utils)

# Read the data
covid <- read.csv("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv", na.strings = "", fileEncoding = "UTF-8-BOM")
```

## Inspect the Data
```{r}
head(covid)
colnames(covid)
str(covid)
summary(covid)
```

## P-hacking

Let's use the continent categorical data to see if it has significant impact on the covid-19 cases (for the 14 day period)

## Testing significance of Whether the country is in the America continent
```{r continent America}
#library(qvalue)
data <- covid
unique(data$dateRep)
dates<-unique(data$dateRep) #261 days
length(dates)

# make a loop through all the dates
pvals_df <- data.frame(dateRep=character(),
                 pval=double(),
                 pval.adjust=double(),
                 stringsAsFactors=FALSE)
counter <- 0

for(i in 1:length(dates)){
  date<-dates[i]
  print(date)
  data_t <- data[data$dateRep==date,]
  model_t <-lm(Cumulative_number_for_14_days_of_COVID.19_cases_per_100000 ~ continentExp,data=data_t)
  pval <- summary(model_t)$coefficients[-1,4]['continentExpAmerica']
  pval.adjust <- p.adjust(pval, method='fdr', n=length(dates))
  pvals_df <- rbind(pvals_df, c(date,pval,pval.adjust))
  counter <- counter +1
}

colnames(pvals_df) <- c('dateRep','pval','qval')
head(pvals_df)

```

Show significant dates when continent correlates with number of covid19 cases. 
```{r }
# significant dates with p-value (un-adjusted)
pvals_df %>% filter(pval < 0.05)

# significant dates with q-value (fdr correction)
pvals_df %>% filter(qval < 0.05)

# show specific date (significant q-value): 16/09/2020
library(ggplot2)
data1 <- data[data$dateRep=='16/09/2020',]
data1$continentExp
ggplot(data1, aes_string(x='continentExp',y='Cumulative_number_for_14_days_of_COVID.19_cases_per_100000')) + geom_boxplot(aes(fill=continentExp))
```
America looks significant in the boxplot. In this case the relationship could be causal, because America continent had times of peaking outbreaks.

## Testing significance of Whether the country is in the Asia continent
```{r continent Asia}
#library(qvalue)
data <- covid
unique(data$dateRep)
dates<-unique(data$dateRep) #261 days
length(dates)

# make a loop through all the dates
pvals_df <- data.frame(dateRep=character(),
                 pval=double(),
                 pval.adjust=double(),
                 stringsAsFactors=FALSE)
counter <- 0

for(i in 1:length(dates)){
  date<-dates[i]
  print(date)
  data_t <- data[data$dateRep==date,]
  model_t <-lm(Cumulative_number_for_14_days_of_COVID.19_cases_per_100000 ~ continentExp,data=data_t)
  pval <- summary(model_t)$coefficients[-1,4]['continentExpAsia']
  pval.adjust <- p.adjust(pval, method='fdr', n=length(dates))
  pvals_df <- rbind(pvals_df, c(date,pval,pval.adjust))
  counter <- counter +1
}

colnames(pvals_df) <- c('dateRep','pval','qval')
head(pvals_df)

```

Show the significant dates when continent correlates with number of covid19 cases. 
```{r }
# significant dates with p-value (un-adjusted)
pvals_df %>% filter(pval < 0.05)

# significant dates with q-value (fdr correction)
pvals_df %>% filter(qval < 0.05)

# show specific date (significant p value wihtout fdr correction): 30/04/2020
library(ggplot2)
data1 <- data[data$dateRep=='30/04/2020',]
data1$continentExp
ggplot(data1, aes_string(x='continentExp',y='Cumulative_number_for_14_days_of_COVID.19_cases_per_100000')) + geom_boxplot(aes(fill=continentExp))

```

### (1) Results:

In the Asia continent, continent factor had spurious significant correlations in 116 dates (without multiple testing adjustments). We know these dates are spurious because after FDR correction whether the country is in the Asia continent did not show any significant correlation to the number of covid-19 cases.

### (2) Non-statistical argument or explanation:
(Argue for an underlying non-statistical explanation for your group’s fake result.)
In some dates, countries in Asia might have had simultaneous spikes at the same day. Since the days that have significant p-value are all after April, these dates are pretty far from when covid-19 outbroke in China.

### (3) Critique your statistical approach and why your group got an apparently significant p-value.

When using the continent as an independent variable, it would be more accurate to use the p-value of the whole model (not p-value for individual coefficients) in order to test the effect of the continent. 88 dates show significant p-value, and among them 47 dates show significant q-value (FDR corrected p-value). Therefore, we could conclude that even though there are some spurious correlations found due to multiple testing, about half of these 88 dates turns out to be significant. 
```{r}

# extract p-value
# function to extract p-value from lm()
lmp <- function (modelobject) {
    if (class(modelobject) != "lm") stop("Not an object of class 'lm' ")
    f <- summary(modelobject)$fstatistic
    p <- pf(f[1],f[2],f[3],lower.tail=F)
    attributes(p) <- NULL
    return(p)
}

#library(qvalue)
data <- covid
unique(data$dateRep)
dates<-unique(data$dateRep) #261 days
length(dates)

# make a loop through all the dates
pvals_df <- data.frame(dateRep=character(),
                 pval=double(),
                 pval.adjust=double(),
                 stringsAsFactors=FALSE)
counter <- 0

for(i in 1:length(dates)){
  date<-dates[i]
  print(date)
  data_t <- data[data$dateRep==date,]
  model_t <-lm(Cumulative_number_for_14_days_of_COVID.19_cases_per_100000 ~ continentExp,data=data_t)
  pval <- lmp(model_t)
  pval.adjust <- p.adjust(pval, method='fdr', n=length(dates))
  pvals_df <- rbind(pvals_df, c(date,pval,pval.adjust))
  counter <- counter +1
}

colnames(pvals_df) <- c('dateRep','pval','qval')
head(pvals_df)

```


```{r}
# significant dates with p-value (un-adjusted)
pvals_df %>% filter(pval < 0.05)

# significant dates with q-value (fdr correction)
pvals_df %>% filter(qval < 0.05)

# show specific date (significant p value wihtout fdr correction): 16/09/2020
library(ggplot2)
data1 <- data[data$dateRep=='10/09/2020',]
data1$continentExp
ggplot(data1, aes_string(x='continentExp',y='Cumulative_number_for_14_days_of_COVID.19_cases_per_100000')) + geom_boxplot(aes(fill=continentExp))

```



